An example exploratory analysis script

### 2024-04-04 R/analysis-code/exploratory-analysis.qmd
part 4

```{r}
# linear model of BMI by water and FAF
lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
lm1_result <- summary(lm1)

@@

@@

```{r}
mydata1 <- mydata %>%
  select(History, Water, Alcohol, FAF, MTRANS, BMI)
lm3 <- lm(BMI ~ ., data = mydata1)
lm3_result <- summary(lm3)

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```
# Notes

>>>>

```{r}
# linear model of BMI by water and FAF
lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
lm1_result <- summary(lm1)
lm1_result

@@

I haven't include gender or age for model fitting, I will try to combine them in one model here.

```{r}
lm4 <- lm(BMI ~ Gender * Age, data = mydata)
lm4_result <- summary(lm4)
lm4_result

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```

This linear model proves gender and age can also be predictors for BMI and there is interaction between them. So I will include them in the all predictor model.

@@

```{r}
lm3 <- lm(BMI ~ Gender + Age + History + Water + Alcohol + FAF + MTRANS, data = mydata)
lm3_result <- summary(lm3)
lm3_result

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```

Choose predictors only important for the model fitting. Thanks to my classmates' replies, I received some useful suggestion for choosing predictors, which are chi-square, LASSO and lowest RMSE. 
According to my searching result online, chi-square is good for categorical variable as outcome, and LASSO might shrink the coefficients too much. So I will go with the lowest RMSE in the following and see what will happen next.

```{r}
# Load necessary packages
library(caret)

# Split data into training and testing sets
set.seed(123)  # Set seed for reproducibility
trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)
trainData <- mydata[trainIndex, ]
testData <- mydata[-trainIndex, ]

# find important predictors without interactions

# Define predictors
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")

# Initialize variables to store results
minRMSE1 <- Inf
bestSubset1 <- NULL

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors)) {
  subsets <- combn(predictors, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE1 < minRMSE1) {
      minRMSE1 <- RMSE1
      bestSubset1 <- predictors_subset
    }
  }
}

# Print results
print(paste("Best subset of predictors:", paste(bestSubset1, collapse = ", ")))
print(paste("Lowest RMSE:", minRMSE1))
```

```{r}
# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
interaction_terms <- c("Gender_Age", "FAF_Water")  # Define interaction terms
response <- "BMI"

# Create interaction terms in training and testing data
trainData$Gender_Age <- trainData$Gender * trainData$Age
trainData$FAF_Water <- trainData$FAF * trainData$Water

testData$Gender_Age <- testData$Gender * testData$Age
testData$FAF_Water <- testData$FAF * testData$Water

# Update list of predictors to include interaction terms
predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")

# Initialize variables to store results
minRMSE2 <- Inf
bestSubset2 <- NULL

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors_updated)) {
  subsets <- combn(predictors_updated, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE2 < minRMSE2) {
      minRMSE2 <- RMSE2
      bestSubset2 <- predictors_subset
    }
  }
}

# Print results
print(paste("Best subset of predictors (including interactions):", paste(bestSubset2, collapse = ", ")))
print(paste("Lowest RMSE (including interactions):", minRMSE2))
```

It is under my expectation that all-predictor model with the 2 interactions has the lowest RMSE. So I also would like to try LASSO regression model here and see the result.

```{r}
# Load necessary packages
library(glmnet)
library(caret)

# Set seed for reproducibility
set.seed(123)

# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
response <- "BMI"

# Prepare data
x <- model.matrix(as.formula(paste(response, "~", paste(predictors, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
y <- mydata[, response]

# Fit LASSO regression model using cross-validation
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- coef(lasso_model, s = best_lambda)

# Identify non-zero coefficients (selected predictors)
selected_predictors1 <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
```

```{r}
# Fit LASSO regression model using cross-validation with interaction terms
x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
y <- mydata[, response]
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- as.matrix(coef(lasso_model, s = "lambda.min"))

# Identify non-zero coefficients (selected predictors)
selected_predictors <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
```

The LASSO regression result includes some specific levels in variables instead of the entire variables.

Using CV to test evaluate the model with all predictors.

```{r}
# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
interaction_terms <- c("Gender_Age", "FAF_Water")
response <- "BMI"

# Define training control with repeated k-fold cross-validation
train_control <- trainControl(method = "repeatedcv", 
                              number = 10,       # Number of folds
                              repeats = 3)       # Number of repetitions

# Define the model formula
formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))

# Train the model using cross-validation
model <- train(formula,
               data = mydata,
               method = "lm",                    # Use linear regression
               trControl = train_control,       # Use defined training control
               preProcess = c("center", "scale"),  # Preprocess data
               metric = "RMSE")                    # Use RMSE as the metric

# Print the cross-validation results
print(model)
```

# Notes
